---
title: "Kmeans Clustering Practice Part 1"
output: html_document
date: "2025-03-24"
---

Today, we're going to do some basic practice with kmeans. Run the chunk below to load the `tidyverse` and read in some sample data.

```{r}
library(tidyverse)
df = read.csv("https://raw.githubusercontent.com/kesslerz-dickinson/Data180_Fall24/refs/heads/main/data/cluster_data.csv")
```

# Exercise 1:
Get a feel for the data. Create a visualization to see if the data are a good candidate for clustering.
```{r}

```

# Exercise 2:
Let's run the kmeans clustering algorithm with `k=3` first. Be sure to save the output to a variable. Then return the centers and include them in a visualization like the one above. Let's also color the original observations by the cluster they are a member of right now.
```{r}

```
# Exercise 3:
This fit is clearly pretty bad (Note we can see how bad by looking at `result$tot.withinss`). Let's repeat what we did in Exercise 2 but with 5 clusters. Run this code a few time. What is happening? 
```{r}

```

# Exercise 4: 
Let's try to find a better fit using the `nstart` parameter in the `kmeans` function. Set `nstart=50` and then recreate the outputs from above again. Now what happens when you run the chunk a few times?"
```{r}

```

