---
title: "DATA 180 Exam 2 - Practice 1"
subtitle: "Spring 2025, Professor Kessler"
output: 
  html_document
---

## Problem 1

In this problem, we will work with the mtcars dataset, which contains specifications and performance measurements of various car models. The dataset includes the following variables:

mpg: miles per gallon
cyl: number of cylinders
disp: displacement (cu. in.)
hp: gross horsepower
drat: rear axle ratio
wt: weight (1000 lbs)
qsec: 1/4 mile time
vs: engine (0 = V-shaped, 1 = straight)
am: transmission (0 = automatic, 1 = manual)
gear: number of forward gears
carb: number of carburetors

Assume the dataset is loaded into an object named `cars`.
```{r}
library(tidyverse)
cars = mtcars
```

a) Use dplyr to compute which number of cylinders (cyl) is associated with the highest average miles per gallon.
```{r}

```

b) Use dplyr to compute how many cars have more than 150 horsepower but weigh less than 3.0 (in 1000 lbs).
```{r}

```

c) Create a new variable called power_to_weight equal to horsepower divided by weight. Then compute the average power_to_weight for automatic vs. manual cars (am).
```{r}

```

d) Run the chunk below to create subset of the cars dataset with only the columns: cyl, am, hp, wt, and gear.

```{r}
cars_small <- cars %>% select(cyl,am,hp,wt,gear)
```

Calculate a similarity matrix using car_small. Be sure to convert categorical variables (or discretes that are effectively categoricals) appropriately. Print the top 6 rows and columns of the resulting matrix.
```{r}
library(cluster)
```


## Problem 2
Consider the following distance matrix:

$$
D = \begin{bmatrix} 
0 & 3.5 & 4.4 & 1.9 & 2.6 \\
3.5 & 0 & 2.7 & 6.1 & 3.3 \\
4.4 & 2.7 & 0 & 4.3 & 5.0 \\
1.9 & 6.1 & 4.3 & 0 & 4.9 \\
2.6 & 3.3 & 5.0 & 4.9 & 0
\end{bmatrix}
$$

Compute the following:

a) Single linkage distance between A = \{O1, O4\} and B = \{O2, O3\}.

b) Complete linkage distance between Group C = \{O2, O5\} and Group D = \{O1, O3\}

c) Average linkage distance between Group E = \{O3, O4\} and Group F = \{O2, O5\}


## Problem 3

Let the following be a 3-by-6 matrix:

$$
M = \begin{bmatrix}
6 & 9 & 11 & 13 & 2 & 5 \\
5 & 3 & 12 & 6 & 8 & 9 \\
10 & 4 & 7 & 2 & 14 & 6
\end{bmatrix}
$$

a) Compute the Manhattan distance between Object 1 and Object 2.



b) Create and print the distance matrix in R. Include the diagonal, but do not include the upper half of the matrix.
```{r}
M <- matrix(c(
  6,  9, 11, 13,  2,  5,
  5,  3, 12,  6,  8,  9,
 10,  4,  7,  2, 14,  6
), nrow = 3, byrow = TRUE)

```


## Problem 4

Use the built-in `USArrests` dataset and perform k-means clustering with `k = 3`.

```{r}
data("USArrests")
scaled_data <- scale(USArrests)
set.seed(42)
kmeans_result <- kmeans(scaled_data, centers = 3)
```

a) Report the size of each cluster.
```{r}

```

b) Which states are assigned to the same cluster as New York?
```{r}

```

c) Use a method to identify the optimal number of clusters for this data.
```{r}

```


## Problem 5
Run the chunk below to import the book, Dracula, as well as some necessary libraries.
```{r}
text <- scan('Dracula.txt', what = "", sep = "\n")
library(quanteda.textstats)
library(quanteda.textplots)
library(quanteda)

char_vec = tolower(text)
```

a) Tokenize the text, remove stopwords, symbols, and punctuation and convert to a term document matrix. What are the 10 most common words?
```{r}

```


b) Update your tokens such that they now are bigrams. Create a wordcloud with these tokens.
```{r}

```


## Problem 6

Consider the following matrix:
```{r}
H <- matrix(c(
  5,  3,  6,
  4,  8,  2,
  9,  1,  7,
  6,  2, 10,
  3,  5,  4
), nrow = 5, byrow = TRUE)
```

a) Perform hierarchical clustering on the matrix H using Euclidean distance and complete linkage. Assign the result to an object called hc.
```{r}

```


b) Create a dendrogram from the result in part (a).
```{r}

```

c) Use the cutree() function to divide the data into 2 clusters. Report the cluster membership for each observation.
```{r}

```

## Short Answer

1. What does cutree() do to a hierarchical clustering result?

2. What kind of variables can daisy() handle that dist() cannot?

3. What does kmeans() return in its $cluster output?

4. What does as.dist() do to a matrix?

5. What type of object does hclust() take as input?

6. What is the effect of increasing k in k-means clustering?

7. What does mutate() do in dplyr?

8. What does tokens() do in quanteda?

9. What is the one distance metric we discussed which uses a parameter value?

10. What is the function we call in summarize() to retrieve the count of observations?

11. What is the name of metric used to evaluate the performance of a kmeans algorithm?

12. How do we force a kmeans algorithim to run multiple times and save the best result?

13. What does the rename() function in dplyr do?

14. How do I convert a similarity matrix to a dissimilarity matrix? 

15. What was it called when we stripped a set of words in a document to only their frequencies and ignored any parts of speech/references/other information? 

---

