---
title: "DATA 180 Exam 2 - Practice 1"
subtitle: "Spring 2025, Professor Kessler"
output: 
  html_document
---

## Problem 1

In this problem, we will work with the mtcars dataset, which contains specifications and performance measurements of various car models. The dataset includes the following variables:

mpg: miles per gallon
cyl: number of cylinders
disp: displacement (cu. in.)
hp: gross horsepower
drat: rear axle ratio
wt: weight (1000 lbs)
qsec: 1/4 mile time
vs: engine (0 = V-shaped, 1 = straight)
am: transmission (0 = automatic, 1 = manual)
gear: number of forward gears
carb: number of carburetors

Assume the dataset is loaded into an object named `cars`.
```{r}
library(tidyverse)
cars <- mtcars
```

a) Use dplyr to compute which number of cylinders (cyl) is associated with the highest average miles per gallon.
```{r}
cars %>% group_by(cyl) %>% summarize(avg = mean(mpg))
```

b) Use dplyr to compute how many cars have more than 150 horsepower but weigh less than 3.0 (in 1000 lbs).
```{r}
cars %>% filter(hp>150,wt<3)
```

c) Create a new variable called power_to_weight equal to horsepower divided by weight. Then compute the average power_to_weight for automatic vs. manual cars (am).
```{r}
cars %>% mutate(power_to_weight=hp/wt) %>% group_by(am) %>% summarize(avg_ptw=mean(power_to_weight))
```

d) Run the chunk below to create subset of the cars dataset with only the columns: cyl, am, hp, wt, and gear.
```{r}
cars_small <- cars %>% select(cyl,am,hp,wt,gear)
```

Calculate a similarity matrix using car_small. Be sure to convert categorical variables (or discretes that are effectively categoricals) appropriately. Print the top 6 rows and columns of the resulting matrix.
```{r}
library(cluster)
output <- cars_small %>% mutate(cyl=as.factor(cyl),am=as.factor(am),gear=as.factor(gear)) %>% daisy(type=list(symm='am')) %>% as.matrix()
output = 1-output
head(output)
```


## Problem 2
Consider the following distance matrix:

$$
D = \begin{bmatrix} 
0 & 3.5 & 4.4 & 1.9 & 2.6 \\
3.5 & 0 & 2.7 & 6.1 & 3.3 \\
4.4 & 2.7 & 0 & 4.3 & 5.0 \\
1.9 & 6.1 & 4.3 & 0 & 4.9 \\
2.6 & 3.3 & 5.0 & 4.9 & 0
\end{bmatrix}
$$

Compute the following:

a) Single linkage distance between A = \{O1, O4\} and B = \{O2, O3\}.

min(3.5, 4.4, 6.1, 4.3) = 3.5

b) Complete linkage distance between Group C = \{O2, O5\} and Group D = \{O1, O3\}

max(3.5, 2.7, 2.6, 5.0) = 5.0

c) Average linkage distance between Group E = \{O3, O4\} and Group F = \{O2, O5\}

sum(2.7, 5.0, 6.1, 4.9) / (2*2) = 4.675


## Problem 3

Let the following be a 3-by-6 matrix:

$$
M = \begin{bmatrix}
6 & 9 & 11 & 13 & 2 & 5 \\
5 & 3 & 12 & 6 & 8 & 9 \\
10 & 4 & 7 & 2 & 14 & 6
\end{bmatrix}
$$

a) Compute the Manhattan distance between Object 1 and Object 2.

|6-5|+|9-3|+|11-12|+|13-6|+|2-8|+|5-9| = 25

b) Create and print the distance matrix in R. Include the diagonal, but do not include the upper half of the matrix.
```{r}
M <- matrix(c(
  6,  9, 11, 13,  2,  5,
  5,  3, 12,  6,  8,  9,
 10,  4,  7,  2, 14,  6
), nrow = 3, byrow = TRUE)

dist(M,method='manhattan',diag=T,upper=F)
```


## Problem 4

Use the built-in `USArrests` dataset and perform k-means clustering with `k = 3`.

```{r}
data("USArrests")
scaled_data <- scale(USArrests)
set.seed(42)
kmeans_result <- kmeans(scaled_data, centers = 3)
```

a) Report the size of each cluster.
```{r}
kmeans_result$size
```

b) Which states are assigned to the same cluster as New York?
```{r}
USArrests$cluster = kmeans_result$cluster
USArrests %>% filter(cluster==2)
```

c) Use a method to identify the optimal number of clusters for this data.
```{r}
wgss<-rep(0,8)
for(i in 1:8){
wgss[i]<-kmeans(scaled_data,centers=i,nstart=50)$tot.withinss
}
ggplot(as.data.frame(wgss), aes(x=1:8,y=wgss)) + geom_point(size=2) + geom_line(linetype = "dashed")
```


## Problem 5
Run the chunk below to import the book, Dracula, as well as some necessary libraries.
```{r}
text <- scan('Dracula.txt', what = "", sep = "\n")
library(quanteda.textstats)
library(quanteda.textplots)
library(quanteda)

char_vec = tolower(text)
```

a) Tokenize the text, remove stopwords, symbols, and punctuation and convert to a term document matrix. What are the 10 most common words?
```{r}
tdm <- char_vec %>% tokens(remove_punct=T,remove_symbols=T,remove_numbers=T) %>% tokens_remove(stopwords('en')) %>% dfm()
textstat_frequency(tdm,n=10)
```


b) Update your tokens such that they now are bigrams. Create a wordcloud with these tokens.
```{r}
tdm <- char_vec %>% tokens(remove_punct=T,remove_symbols=T,remove_numbers=T) %>% tokens_remove(stopwords('en')) %>% tokens_ngrams(n=2) %>% dfm()
textplot_wordcloud(tdm)
```


## Problem 6

Consider the following matrix:
```{r}
H <- matrix(c(
  5,  3,  6,
  4,  8,  2,
  9,  1,  7,
  6,  2, 10,
  3,  5,  4
), nrow = 5, byrow = TRUE)
```

a) Perform hierarchical clustering on the matrix H using Euclidean distance and complete linkage. Assign the result to an object called hc.
```{r}
d <- dist(H)
hc <- hclust(d)
```


b) Create a dendrogram from the result in part (a).
```{r}
plot(hc)
```

c) Use the cutree() function to divide the data into 2 clusters. Report the cluster membership for each observation.
```{r}
memberships <- cutree(hc,k=2)
#O1, O2, and O5 are in Cluster 1. O3 and O4 are in Cluster 2.
```

## Short Answer

1. What does cutree() do to a hierarchical clustering result?

Retrieve membership of observations at a specific tree level.

2. What kind of variables can daisy() handle that dist() cannot?

Categorical

3. What does kmeans() return in its $cluster output?

Cluster Memberships

4. What does as.dist() do to a matrix?

Convert matrix to a distance object

5. What type of object does hclust() take as input?

a distance object

6. What is the effect of increasing k in k-means clustering?

Add another centroid/increase number of clusters

7. What does mutate() do in dplyr?

Create a new column according to some operation

8. What does tokens() do in quanteda?

Convert a vector or corpus into word tokens.

9. What is the one distance metric we discussed which uses a parameter value?

Minkowski

10. What is the function we call in summarize() to retrieve the count of observations?

n()

11. What is the name of metric used to evaluate the performance of a kmeans algorithm?

Within Group Sum of Squares

12. How do we force a kmeans algorithim to run multiple times and save the best result?

nstart

13. What does the rename() function in dplyr do?

Renames a column

14. How do I convert a similarity matrix to a dissimilarity matrix? 

D = 1-S

15. What was it called when we stripped a set of words in a document to only their frequencies and ignored any parts of speech/references/other information? 

Bag of words

---

