---
title: "R Practice Exam Problems"
subtitle: "Spring 2025"
output: 
  html_document
---

## Problem 1

Suppose we have a vector named `mycharacters` with popular movie characters:

```{r}
mycharacters <- c('Harry Potter', 'Darth Vader', 'Frodo Baggins', 'Indiana Jones', 'Sherlock Holmes')
```

a) Add `Luke Skywalker` as the final item in this vector. Do not rewrite the whole vector.

```{r}
mycharacters[6]='Luke Skywalker'
```

b) Write a loop that iterates through and prints every element in the `mycharacters` vector.

```{r}
for (char in mycharacters){
  print(char)}
```

## Problem 2

Suppose we want to create the following data frame named `movie_earnings`. Write R code that creates the data frame shown below. Earnings is in number of billions.


| Character        | Earnings            | RottenTomatoes       |
|-----------------|----------------------|------------------|
| Harry Potter    | 7.7                  | 84                   |
| Darth Vader     | 8.5                  | 92                   |
| Frodo Baggins   | 8.6                  | 94                   |
| Indiana Jones   | 5.8                  | 95                   |
| Sherlock Holmes | 4.2                  | 87                   |
| Luke Skywalker  | 6.3                  | 89                   |

```{r}
movie_earnings <- data.frame(
  Character = c("Harry Potter", "Darth Vader", "Frodo Baggins", "Indiana Jones", "Sherlock Holmes", "Luke Skywalker"),
  Earnings = c(7.7, 8.5, 8.6, 5.8, 4.2, 6.3),
  RottenTomatoes = c(84, 92, 94, 95, 87, 89)
)
```

## Problem 3: Filtering and Modifying a Data Frame

Using the `movie_earnings` dataset from above, perform the following tasks:

 a) Filter the dataset to return only rows with earnings greater than 7 billion.

```{r}
movie_earnings[movie_earnings$Earnings>7,]
```

 b) Filter the dataset for movies with a Rotten Tomatoes Score greater than 85 and return only the `Character` and `Earnings` columns.

```{r}
movie_earnings[movie_earnings$RottenTomatoes>85,c("Character","Earnings")]
```

 c) Add a new column called `Franchise` that categorizes each character according to the film series they are a part of.

```{r}
movie_earnings$Franchise <- c("Harry Potter", "Star Wars", "Lord of the Rings", "Indiana Jones", "Sherlock Holmes", "Star Wars")
```

## Problem 4: 

 a) Define a function
Write a function named `my_func` that takes a numeric input, multiplies it by 2, and then adds 5 to the result.

```{r}
my_func <- function(x) {
  return((x * 2) + 5)
  }
```

 b) Create a for loop
Write a for loop that runs 5 times, applying `my_func` to the loop index (`i`) in each iteration, and summing all the results.

```{r}
sum_result <- 0
for (i in 1:5) {
  sum_result <- sum_result + my_func(i)
}
sum_result
```

## Problem 5: Base R Visualizations

```{r}
data(PlantGrowth)
```

Using the built-in **PlantGrowth** dataset, create the following visualizations. Label you axis appropriately and add titles.

 a) Barplot
Create a barplot showing the number of relative frequency each treatment group.

```{r}
barplot(table(PlantGrowth$group)/nrow(PlantGrowth),
        main = "Relative Frequency of Treatment Groups",
        xlab = "Treatment Group", 
        ylab = "Relative Frequency",
        col = 4)
```

 b) Histogram
Create a histogram of the plant weights. Make sure it follows the rule of thumb.

```{r}
hist(PlantGrowth$weight, 
     main = "Histogram of Plant Weights", 
     xlab = "Weight", 
     col = 6,
     breaks=5)
```

 c) Pie
 Create a pie chart of the frequency distribution of each treatment group.

```{r}
pie(table(PlantGrowth$group),
    main = "Frequency Distribution of Treatment Groups",
    col = 1:3)
```

## Problem 6: ggplot2 Visualizations

```{r}
library(ggplot2)
data(mtcars)
```

Using the built-in **mtcars** dataset, create the following visualizations:

 a) Barplot
Create a barplot showing the count of cars for each cylinder category.

```{r}
ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar(fill = 8) +
  labs(title = "Count of Cars by Cylinder Category", x = "Cylinders", y = "Count")
```

 b) Histogram
Create a histogram of the `mpg` (miles per gallon) column. Make sure it follows the rule of thumb.

```{r}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(fill = 9,bins=4) +
  labs(title = "Histogram of Miles Per Gallon", x = "MPG", y = "Count")
```

 c) Scatterplot with Smoothing
Create a scatterplot of horsepower (`hp`) vs. weight (`wt`) and overlay a smoothing line.

```{r}
ggplot(mtcars, aes(x = wt, y = hp)) +
  geom_point(color = "red", size = 3,shape=13) +
  geom_smooth(method = "lm",linetype=2) +
  labs(title = "Scatterplot of Horsepower vs. Weight with Smoothing", x = "Weight (1000 lbs)", y = "Horsepower")
```

## Short Answer

1. What is the key difference between a sample and a population?  Portion v distribution

2. If a variable `education_level` is coded such that high school = 1, bachelor's = 2, master's = 3, PhD = 4, what type of variable is `education_level`?  Ordinal

3. Why might the median be preferred over the mean in a dataset with extreme outliers?  Mean is pulled by outliers, median is not

4. A distribution that has a long right tail is called what?  Right-skew

5. When a dataset has 1 mode, we call it what? unimodal

6. If a variable `phone_number` stores a person’s 10-digit phone number, what type of variable is `phone_number`? Nominal

7. What are variance and standard deviation attempting to measure? spread

8. What do we call the process of solving a mathematical problem backwards?  recursion

9. Name one statistic that is considered non-resistant.  mean

10. What is it called when a dataset has more observations on one side, creating a longer tail?  skewness

11. Write a single line of R code that returns the following sequence: (5,10,15,20,25,30). Do not simply write out `c(5,10,15,20,25,30)`.  
    ```{r}  
print(seq(5,30,5))
    ```  

12. What is the key difference between categorical and numerical variables? Expresses info on a category versus the count or measure of something

13. What is the primary difference between a scatterplot and a line plot? line plot connects points with a line scatter places just the points

14. Which of the following R functions is used to read a CSV file? b
    a) `read.table()`  
    b) `read.csv()`  
    c) `readLines()`  
    d) `scan()`  

15. What type of numeric variable might we use a barplot to show the distribution for? Discrete
