---
title: "R Practice Exam Problems"
subtitle: "Spring 2025"
output: 
  html_document
---

## Problem 1

Suppose we have a vector named `mycharacters` with popular movie characters:

```{r}
mycharacters <- c('Harry Potter', 'Darth Vader', 'Frodo Baggins', 'Indiana Jones', 'Sherlock Holmes')
```

a) Add `Luke Skywalker` as the final item in this vector. Do not rewrite the whole vector.

```{r}

```

b) Write a loop that iterates through and prints every element in the `mycharacters` vector.

```{r}

```

## Problem 2

Suppose we want to create the following data frame named `movie_earnings`. Write R code that creates the data frame shown below. Earnings is in number of billions.


| Character        | Earnings            | RottenTomatoes       |
|-----------------|----------------------|----------------------|
| Harry Potter    | 7.7                  | 84                   |
| Darth Vader     | 8.5                  | 92                   |
| Frodo Baggins   | 8.6                  | 94                   |
| Indiana Jones   | 5.8                  | 95                   |
| Sherlock Holmes | 4.2                  | 87                   |
| Luke Skywalker  | 6.3                  | 89                   |

```{r}

```

## Problem 3: Filtering and Modifying a Data Frame

Using the `movie_earnings` dataset from above, perform the following tasks:

 a) Filter the dataset to return only rows with earnings greater than 7 billion.

```{r}

```

 b) Filter the dataset for movies with a Rotten Tomatoes Score greater than 85 and return only the `Character` and `Earnings` columns.

```{r}

```

 c) Add a new column called `Franchise` that categorizes each character according to the film series they are a part of.

```{r}

```

## Problem 4: 

 a) Define a function
Write a function named `my_func` that takes a numeric input, multiplies it by 2, and then adds 5 to the result.

```{r}

```

 b) Create a for loop
Write a for loop that runs 5 times, applying `my_func` to the loop index (`i`) in each iteration, and summing all the results.

```{r}

```

## Problem 5: Base R Visualizations

```{r}
data(PlantGrowth)
```

Using the built-in **PlantGrowth** dataset, create the following visualizations. Label you axis appropriately and add titles.

 a) Barplot
Create a barplot showing the number of relative frequency each treatment group.

```{r}

```

 b) Histogram
Create a histogram of the plant weights. Make sure it follows the rule of thumb.

```{r}

```

 c) Pie
 Create a pie chart of the frequency distribution of each treatment group.

```{r}

```

## Problem 6: ggplot2 Visualizations

```{r}
library(ggplot2)
data(mtcars)
```

Using the built-in **mtcars** dataset, create the following visualizations:

 a) Barplot
Create a barplot showing the count of cars for each cylinder category.

```{r}

```

 b) Histogram
Create a histogram of the `mpg` (miles per gallon) column. Make sure it follows the rule of thumb.

```{r}

```

 c) Scatterplot with Smoothing
Create a scatterplot of horsepower (`hp`) vs. weight (`wt`) and overlay a smoothing line.

```{r}

```

## Short Answer

1. What is the key difference between a sample and a population?  

2. If a variable `education_level` is coded such that high school = 1, bachelor's = 2, master's = 3, PhD = 4, what type of variable is `education_level`? 

3. Why might the median be preferred over the mean in a dataset with extreme outliers?  

4. A distribution that has a long right tail is called what?  

5. When a dataset has 1 mode, we call it what?

6. If a variable `phone_number` stores a person’s 10-digit phone number, what type of variable is `phone_number`?  

7. What are variance and standard deviation attempting to measure?  

8. What do we call the process of solving a mathematical problem backwards?  

9. Name one statistic that is considered non-resistant.  

10. What is it called when a dataset has more observations on one side, creating a longer tail?  

11. Write a single line of R code that returns the following sequence: (5,10,15,20,25,30). Do not simply write out `c(5,10,15,20,25,30)`.  
    ```{r}  

    ```  

12. What is the key difference between categorical and numerical variables?  

13. What is the primary difference between a scatterplot and a line plot?  

14. Which of the following R functions is used to read a CSV file?  
    a) `read.table()`  
    b) `read.csv()`  
    c) `readLines()`  
    d) `scan()`  

15. What type of numeric variable might we use a barplot to show the distribution for?  

