---
title: "R Practice Exam Problems 2"
subtitle: "Spring 2025"
output: html_document
---

## Practice Questions

```{r}
library(datasets)
library(ggplot2)
data(airquality)
airquality=na.omit(airquality)
data(ToothGrowth)
```

## Problem 1 
Using the built-in `airquality` dataset, perform the following tasks:

   a) Filter for rows where `Ozone` levels are greater than 50 and return only the `Ozone` and `Temp` columns.
   
```{r}
 
```
   
   b) Filter the dataset for only July and `Wind` speeds above 9. Return the Ozone column.
   
```{r}
   
```
   
   c) Create a new column `WCI` that converts the `Temp` and `Wind` columns into the Wind Chill Index. This formula is as follows:
   35.74+.6215T-35.75(V^.16)+.4275T(V^.16), where T is the temperature and V is the wind speed.
   
```{r}
   
```
   d) Create a scatterplot using ggplot to  visualize the relationship between `Wind` and your new `WCI` column. Label the axes and describe the relationship.
   
```{r}
   
```

## Problem 2 
Perform the following vector-based tasks:
   
   a) Create a vector named `multiples_of_four` that contains the first 10 multiples of 4 using an efficient method.
   
```{r}
   
```
   
   b) Loop and print out every item from this vector.
   
```{r}
   
```
   
   c) Compute the sum of all elements in `multiples_of_four`.
   
```{r}
   
```

## Problem 3 
Build the following function and implement it in the following tasks:
   
   a) Write a function named `convert_to_celcius()` that takes a temperature in Fahrenheit as an argument and returns the equivalent Celsius temperature using the formula: (5/9)*(T-32), where T is the temperature
   
```{r}
   
```
   
   b) Modify `convert_to_celcius()` to handle a vector of Fahrenheit values as input and return a vector of Celsius values.
   
```{r}
   
```
   
   c) Use `convert_to_celcius()` to convert the `Temp` column from the `airquality` dataset into Celsius. Make a new column `Temp_C`.
   
```{r}
   
```

## Problem 4 
Using the `ToothGrowth` dataset, complete the following visualizations using Base R AND ggplot:

   a) Create a barplot showing the number of observations for each `dose` level.
   
```{r}
   
```
   
   b) Create a histogram comparing `len` distributions across the `dose` levels. Your answer should have multiple histograms. Do this only in ggplot.
   
```{r}
   
```

## Short Answer

1. What is the key difference between continuous and discrete variables?

2. If a variable `income_level` is coded as low = 1, medium = 2, high = 3, what type of variable is `income_level`?

3. If a variable `hot_days` stores a numeric value measured as the count of days which reached a temperature above 90 degrees in a particular month, what type of variable is `hot_days`?

4. When a distribution has relatively equal sized tails, we call it what?

5. When you knit an R markdown file as we have done in homework. What is the resulting file type?

   a) `.Rmd`  
   b) `.html`  
   c) `.git`  
   d) `.csv`  

6. What symbol do we use the specify greater than or equal to when filtering a dataframe?

7. What function in R returns the first few rows of a dataframe? Implement it on an object called `dataset`.

8. Write a single line of R code that returns the following sequence: (3, 6, 9, 12, 15, 18). Do not simply write out `c(3,6,9,12,15,18)`.

```{r}
   
```

9. What is a resistant statistic? Give one example.

10. A dataset has two distinct peaks in its distribution. What is this type of distribution called?

11. What are mean, median, and mode attempting to measure?

12. Why is variance less useful than standard deviation for analysis?

13. What is the primary difference between a statistic and a parameter in statistics?

14. Can a dataset be considered both a population and a sample? Explain with an example.

15. What is the purpose of `geom_smooth()` in `ggplot2`? How does it enhance a scatterplot?
