---
title: "Dataframe Exercies"
subtitle: "Data 180, Professor Kessler"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---

## Introduction

This document contains a set of exercises to help you practice working with data frames in R. You will learn how to create dataframes, count rows and columns, perform basic filtering operations, and calculate some summary statistics.

### First steps...

Create a data frame named `students` with the following columns:
- `Name` (character): "Alice", "Bob", "Charlie", "David", "Eva"
- `Age` (numeric): 18, 21, 20, 19, 18
- `Major` (character): "Math", "Physics", "Biology", "Chemistry", "Math"
- `Starting Year` (numeric):2024,2021,2022,2023,2024

# Exercise 1: Create the data frame

```{r}
students <- data.frame(
  Name = c("Alice", "Bob", "Charlie", "David", "Eva"),
  Age = c(22, 23, 22, 24, 21),
  Major = c("Math", "Physics", "Biology", "Chemistry", "Math"),
  FirstYear = c(2024,2021,2022,2023,2024)
)

students
```

# Exercise 2: Count rows, columns, and dimensions
```{r}
num_rows <- nrow(students)
num_cols <- ncol(students)

num_rows
num_cols

dim(students)
```
# Exercise 3: Filter for Math majors
```{r}
students[students$Major=='Math',]
students[students$Major=='Math',"Age"]
math_students <- students[students$Major=='Math',c("Age","FirstYear")]
math_students

```
# Exercise 4: Filter for students older than 22
```{r}
older_students <- students[students$Age > 22,]
older_students
```
# Exercise 5: Filter for Math majors older than 22
```{r}
math_older_students <- students[students$Major == "Math" & students$Age >20,]
math_older_students
```
# Exercise 6: Add GraduationYear column
```{r}
students$GraduationYear <- students$FirstYear + 4

```
#Exercise 7: Creating a Vector
```{r}
my_vec <- c(1,2,3,4,5)
my_vec
my_vec_2 <- 1:5
my_vec_2
my_vec_3 <- c(1L,2L,3L,4L,5L)
my_vec_3
```
#Exercise 8: Indices
```{r}
my_vec_2[1]<-10
```
#Exercise 9: Creating functions
```{r}
my_function <- function(x,y){
  val1 <- x^2
  val2 <- y+x*y
  result <-val1-val2
  return(result)
}
my_function(5,1)
```
