---
title: "R Practice Exam Problems 2"
subtitle: "Spring 2025"
output: html_document
---

## Practice Questions

```{r}
library(datasets)
library(ggplot2)
data(airquality)
airquality=na.omit(airquality)
data(ToothGrowth)
```

## Problem 1 
Using the built-in `airquality` dataset, perform the following tasks:

   a) Filter for rows where `Ozone` levels are greater than 50 and return only the `Ozone` and `Temp` columns.
   
```{r}
airquality[airquality$Ozone>50,c("Ozone","Temp")]
```
   
   b) Filter the dataset for only July and `Wind` speeds above 9. Return the Ozone column.
   
```{r}
airquality[(airquality$Month==7) & (airquality$Wind>9),"Ozone"]
```
   
   c) Create a new column `WCI` that converts the `Temp` and `Wind` columns into the Wind Chill Index. This formula is as follows:
   35.74+.6215T-35.75(V^.16)+.4275T(V^.16), where T is the temperature and V is the wind speed.
   
```{r}
airquality$WCI <- 35.74 + (0.6215 * airquality$Temp) - (35.75 * (airquality$Wind^0.16)) + (0.4275 * airquality$Temp * (airquality$Wind^0.16))
```
   d) Create a scatterplot using ggplot to  visualize the relationship between `Wind` and your new `WCI` column. Label the axes and describe the relationship. Shape each point by month. Color each point by Ozone.
   
```{r}
ggplot(airquality, aes(x = Wind, y = WCI,shape=as.factor(Month),color=Ozone)) +
  geom_point() +
  labs(title = "Wind vs. Wind Chill Index", x = "Wind Speed", y = "Wind Chill Index")

```

## Problem 2 
Perform the following vector-based tasks:
   
   a) Create a vector named `multiples_of_four` that contains the first 10 multiples of 4 using an efficient method.
   
```{r}
multiples_of_four <- seq(4,40,4)
```
   
   b) Loop and print out every item from this vector.
   
```{r}
for (i in multiples_of_four){
   print(i)
}
```
   
   c) Compute the sum of all elements in `multiples_of_four`.
   
```{r}
sum(multiples_of_four)
```

## Problem 3 
Build the following function and implement it in the following tasks:
   
   a) Write a function named `convert_to_celcius()` that takes a temperature in Fahrenheit as an argument and returns the equivalent Celsius temperature using the formula: (5/9)*(T-32), where T is the temperature
   
```{r}
convert_to_celcius <- function(T) {
  return((5/9) * (T - 32))
}
convert_to_celcius(88)
```
   
   b) Modify `convert_to_celcius()` to handle a vector of Fahrenheit values as input and return a vector of Celsius values.
   
```{r}
   convert_to_celcius <- function(T) {
      C = c()
      for (i in 1:length(T)){
         C[i]=(5/9) * (T[i] - 32)
      }
  return(C)
   }
test=c(93,45,82)
convert_to_celcius(test)
```
   
   c) Use `convert_to_celcius()` to convert the `Temp` column from the `airquality` dataset into Celsius. Make a new column `Temp_C`.
   
```{r}
airquality$Temp_C = convert_to_celcius(airquality$Temp)
```

## Problem 4 
Using the `ToothGrowth` dataset, complete the following visualizations:

   a) Create a barplot showing the number of observations for each `dose` level.Do this in Base R and ggplot.
   
```{r}
barplot(table(ToothGrowth$dose), main = "Number of Observations per Dose Level", xlab = "Dose", ylab = "Count", col = 1)


ggplot(ToothGrowth, aes(x = dose)) +
  geom_bar(fill = 2) +
  labs(title = "Number of Observations per Dose Level", x = "Dose", y = "Count")
```
   
   b) Create a histogram comparing `len` distributions across the `dose` levels. Your answer should have multiple histograms.Do this in ggplot only.
   
```{r}
ggplot(ToothGrowth, aes(x = len)) +
  geom_histogram() +
  facet_wrap(~dose)+
  labs(title = "Tooth Length Distribution by Dose Level", x = "Tooth Length", y = "Count")
```

## Short Answer

1. What is the key difference between continuous and discrete variables? Intgers v values bewteen the integers

2. If a variable `income_level` is coded as low = 1, medium = 2, high = 3, what type of variable is `income_level`? Oridnal

3. If a variable `hot_days` stores a numeric value measured as the count of days which reached a temperature above 90 degrees in a particular month, what type of variable is `hot_days`? Discrete

4. When a distribution has relatively equal sized tails, we call it what? Symmetric

5. When you knit an R markdown file as we have done in homework. What is the resulting file type? b

   a) `.Rmd`  
   b) `.html`  
   c) `.git`  
   d) `.csv`  

6. What symbol do we use the specify greater than or equal to when filtering a dataframe? >=

7. What function in R returns the first few rows of a dataframe? Implement it on an object called `dataset`.

head(dataset)

8. Write a single line of R code that returns the following sequence: (3, 6, 9, 12, 15, 18). Do not simply write out `c(3,6,9,12,15,18)`.

```{r}
seq(3,18,3)
```

9. What is a resistant statistic? Give one example. a stat which resists extreme values/outliers, median

10. A dataset has two distinct peaks in its distribution. What is this type of distribution called? bimodal

11. What are mean, median, and mode attempting to measure? center/centrality

12. Why is variance less useful than standard deviation for analysis? Not in the units of the observations

13. What is the primary difference between a statistic and a parameter in statistics? Statistic for samples, parameter for populations

14. Can a dataset be considered both a population and a sample? Explain with an example. Yes, depends on the reference point for analysis, Class is a sample of Dickinson, population of itself

15. What is the purpose of `geom_smooth()` in `ggplot2`? How does it enhance a scatterplot? Adds a fitted line to the data by building a model, gives us an idea of the relationship between two values.
