---
title: "Debugging and Error Exercises"
output: html_document
---

```{r}
library(MASS)
library(ggplot2)
```


# Exercise 1: Create the data frame

```{r}
students <- data.frame(
  Name = c("Alice Bob Charlie, David, Eva"),
  Age = c(22, 23, 22, 24, 21)
  Major = ("Math", "Physics", "Biology", "Chemistry", "Math"),
  FirstYear = c(2024,2021,2022,2023,2024),
)

students
```

# Exercise 2: Count rows, columns, and dimensions
```{r}
num_rows <- nrow(students,x=1)
num_cols <- ncol(students)

num_rows
num_cols

dimstudents)
```
# Exercise 3: Filter for Math majors
```{r}
students[students$Major=='Math']
students[Major=='Math',"Age"]
math_students <- students[students$Major=='Math',"Age","FirstYear"]
math_students

```
# Exercise 4: Filter for students older than 22
```{r}
older_students <- students[students$Age > 22,]
older_students
```
# Exercise 5: Filter for Math majors older than 22
```{r}
math_older_students <- students[students$Major == Math & students$Age >20,]
math_older_students
```

#Exercise 7: Creating a Vector
```{r}
my_vec <- (1,2,3,4,5)
my_vec

```
#Exercise 9: Creating functions
```{r}
my_function <- function(x,y)
  val1 <- x^2
  val2 <- y+x*y
  result <-val1-val2
  return(result)

my_function(5,1)
```

# Exercise 1: Frequency Distribution with Categorical Variables

## 1.1 Frequency Distribution for Iris Species

```{r}
iris_freq <- table(iris,Species)
iris_freq
```

## 1.2 Frequency Distribution for Car Types (Cars93)

```{r}
data(Cars93)
cars93_freq <- table(c(Type))
cars93_freq
```

## 1.3 Frequency Distribution for Diamond Cuts

```{r}
data(diamonds)
diamonds_freq <- table(diamonds)
diamonds_freq
```

# Exercise 2: Relative Frequency Distributions

## 2.1 Relative Frequency for Iris Species

```{r}
iris_rel_freq <- table(iris$Species)/nrow(iris$Species)
iris_rel_freq
```
# Exercise 3: Barplots

## 3.1 Basic Barplot for Iris Species

```{r}
barplot(table(iris$Species), 
        main = "Frequency of Iris Species", 
        labs = "Species","Frequency", 
        col = "lightblue")
```

## 3.2 Side-by-Side Barplot for Car Types in Cars93

```{r}
cars93_table <- table(Cars93$Type,Cars93$DriveTrain)
cars93_table

barplot(cars93_table, 
        beside = T, 
        main = "Car Types by DriveTrain", 
        xlab = "Car Type", 
        ylab = "Frequency", 
        col = 1,2,3,4,5,6)

```

# Exercise 4: Pie Charts

## 4.1 Pie Chart for Diamond Cuts

```{r}
pie(table(diamonds$cut), 
    main = "Pie Chart of Diamond Cuts", 
    col = rainbow(length(table(diamonds$cut)))
```

# Exercise 5: Histograms

## 5.1 Histogram for Iris Sepal Length

```{r}
hist(iris$Sepal.Length, 
     main = "Histogram of Iris Sepal Length" 
     xlabs = "Sepal Length", 
     cols = "lightgreen", 
     border = "white")
```

