---
title: "Distance Exercises"
output: 
  html_document
---

## Exercise 1:

Given the following two observations:

- Observation A: (2, 5)
- Observation B: (7, 9)

1. Calculate the Euclidean distance between Observation A and Observation B by hand.
2. Run code which does this calculation.

```{r}
mat <- matrix(c(2,7,5,9),nrow=2,byrow=F)
dist(mat)
calc1 <- (mat[1,1]-mat[2,1])^2+(mat[1,2]-mat[2,2])^2
calc2 <- sqrt(calc1)
calc2
```

## Exercise 2: Manhattan Distance

Using the same observations from Exercise 1:

**Tasks:**

1. Calculate the Manhattan distance between Observation A and Observation B by hand.
2. Do this in code.

```{r}
mat <- matrix(c(2,7,5,9),nrow=2,byrow=F)
dist(mat,method='manhattan')
calc11 <-  abs(mat[1,1]-mat[2,1])+abs(mat[1,2]-mat[2,2])
calc11
```


## Exercise 3: Minkowski Distance

Again, using the same observations:


1. Calculate the Minkowski distance between Observation A and Observation B with \( p = 3 \).
2. Do this in code.

```{r}
mat <- matrix(c(2,7,5,9),nrow=2,byrow=F)
dist(mat,method='minkowski',p=3)
p=3
calc111 <-(abs(mat[1,1]-mat[2,1]))^p+(abs(mat[1,2]-mat[2,2]))^p
calc222 <- calc111^(1/p)
calc222
```

## Exercise 4: Calculating Distances in `Cars93`

Load the `MASS` package to access the `Cars93` dataset.

```{r}
library(MASS)
library(tidyverse)
data("Cars93")
```

1. Make a new dataframe with the following numerical variables: `Horsepower`, `MPG.city`, and `EngineSize`. Use only the first 10 rows.
3. Compute the Euclidean distance matrix for these variables using R.
4. Compute the Manhattan distance matrix for these variables using R.
5. Compute the Minkowski distance matrix with \( p = 3 \) using R.
6. Interpret the results: Which cars are most similar according to each distance measure? Does this change depending on the measure?

```{r}
data_new <- Cars93 %>% select(Horsepower, MPG.city,EngineSize) %>% head(10)
dist1 <- dist(data_new)
dist2 <- dist(data_new,method='manhattan')
dist3 <- dist(data_new,method='minkowski',p=3)

dist1
dist2
dist3

```

---

