---
title: "ggplot Introduction Exercises"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(MASS)
library(ggplot2)
data(iris)
data(diamonds)
data(Cars93)
data(txhousing)
```

# Exercise 1: Barplots

## 1.1 Basic Barplot for Iris Species

```{r}
barplot(table(iris$Species), 
        main = "Frequency of Iris Species", 
        xlab = "Species", 
        ylab = "Frequency", 
        col = "lightblue")

```

## Do the same with ggplot:
```{r}
ggplot(iris,aes(x=Species))+
        geom_bar(fill=2)+
        labs(title="Frequency of Iris Species",x="Species",y='Frequency')

```



## 1.2 Side-by-Side Barplot for Car Types in Cars93

```{r}
cars93_table <- table(Cars93$Type, Cars93$DriveTrain)
cars93_table

barplot(cars93_table, 
        beside = T, 
        main = "Car Types by DriveTrain", 
        xlab = "Car Type", 
        ylab = "Frequency", 
        col = 1:6)

```
## Do the same with ggplot:
```{r}
ggplot(Cars93,aes(x=Type,fill=DriveTrain))+
        geom_bar(position='dodge')+
        labs(title='Cars Type by Drive Train',x='Car Type',y='Frequency')

```
# Exercise 2: Histograms

## 2.1 Histogram for Iris Sepal Length

```{r}
hist(iris$Sepal.Length, 
     main = "Histogram of Iris Sepal Length", 
     xlab = "Sepal Length", 
     col = "lightgreen", 
     border = "white")
```
## Do the same with ggplot:
```{r}
ggplot(iris,aes(x=Sepal.Length))+
        geom_histogram(fill=5,bins=round(sqrt(nrow(Cars93))))
```
# Exercise 3: Scatterplot

## 3.1 Scatterplot of diamonds carat and price

```{r}
plot(diamonds$price,diamonds$carat,xlab='Price',ylab='Carat')
```
## Do the same with ggplot but add a smoothing function (method as lm). Also try it without a method included and change the `linetype` to dotted
```{r}
ggplot(diamonds,aes(y=price,x=carat))+
        geom_point(shape=1)+
        geom_smooth(method='lm',color=2,se=FALSE,linetype=3)

ggplot(diamonds,aes(y=price,x=carat))+
        geom_point(shape=1)+
        geom_smooth(color=3,se=FALSE,linetype=3)
        
```
# Exercise 4: Advanced ggplot

## Exercise 4.1 Iris Scatterplot (x is Sepal.Length, y is Sepal.Width) with Point Color determined by Species. Do the same for point shape. How about both?
```{r}
ggplot(iris,aes(Sepal.Length,Sepal.Width,shape=Species,color=Species))+
  geom_point()
```
## Exercise 4.2 Cars93 Histogram using Price for every cyclinder (facet_wrap)
```{r}
ggplot(Cars93,aes(x=Price))+
        geom_histogram(fill=1)+
        facet_wrap(~Cylinders)
```
## Exercise 4.3 Cars93 Histrogram using Horsepower for every type and Drive Train (facet_grid)
```{r}
ggplot(Cars93,aes(x=Horsepower))+
        geom_histogram()+
        facet_grid(rows=vars(Type),cols=vars(DriveTrain))
```
## Exercise 4.4 Scatterplot with Diamonds between depth and table where color is determined by price. Additionally break charts out by cut.
```{r}
ggplot(diamonds,aes(x=depth,y=table,color=price))+
        geom_point()+
        facet_wrap(~cut)
```
## Exercise 4.5 Movement of median price over time in txhousing. Color equal to city.
```{r}
ggplot(txhousing,aes(y=median,x=date,color=city))+
        geom_line(alpha=.5,show.legend=F)
```
## Exercise 4.6 Recreate the Mean v Median Scatterplot from the Abstract Question. Try to identify if any particular cities have an outlier problem. Visualize this in some way.
```{r}
txhousing$mean = txhousing$volume/txhousing$sales

data = txhousing[txhousing$year==2015,]

ggplot(data,aes(median,mean,color=city))+
        geom_point(alpha=.5,show.legend=FALSE)+
        geom_abline()
```
## Exercise 4.7 Now do this for all years, not just 2015
```{r}
ggplot(txhousing,aes(median,mean,color=city))+
        geom_point(show.legend=F,alpha=.5)+
        facet_wrap(~as.factor(year))
```




