---
title: "ggplot Introduction Exercises"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(MASS)
library(ggplot2)
data(iris)
data(diamonds)
data(Cars93)
data(txhousing)
```

# Exercise 1: Barplots

## 1.1 Basic Barplot for Iris Species

```{r}
barplot(table(iris$Species), 
        main = "Frequency of Iris Species", 
        xlab = "Species", 
        ylab = "Frequency", 
        col = "lightblue")

```

## Do the same with ggplot:
```{r}

```



## 1.2 Side-by-Side Barplot for Car Types in Cars93

```{r}
cars93_table <- table(Cars93$Type, Cars93$DriveTrain)
cars93_table

barplot(cars93_table, 
        beside = T, 
        main = "Car Types by DriveTrain", 
        xlab = "Car Type", 
        ylab = "Frequency", 
        col = 1:6)

```
## Do the same with ggplot:
```{r}

```
# Exercise 2: Histograms

## 2.1 Histogram for Iris Sepal Length

```{r}
hist(iris$Sepal.Length, 
     main = "Histogram of Iris Sepal Length", 
     xlab = "Sepal Length", 
     col = "lightgreen", 
     border = "white")
```
## Do the same with ggplot:
```{r}

```
# Exercise 3: Scatterplot

## 3.1 Scatterplot of diamonds carat and price

```{r}
plot(diamonds$price,diamonds$carat,xlab='Price',ylab='Carat')
```
## Do the same with ggplot but add a smoothing function (method as lm). Also try it without a method included and change the `linetype` to dotted
```{r}

```
# Exercise 4: Advanced ggplot

## Exercise 4.1 Iris Scatterplot (x is Sepal.Length, y is Sepal.Width) with Point Color determined by Species. Do the same for point shape. How about both?
```{r}

```
## Exercise 4.2 Cars93 Histogram using Price for every cyclinder (facet_wrap)
```{r}

```
## Exercise 4.3 Cars93 Histrogram using Horsepower for every type and Drive Train (facet_grid)
```{r}

```
## Exercise 4.4 Scatterplot with Diamonds between depth and table where color is determined by price. Additionally break charts out by cut.
```{r}

```
## Exercise 4.5 Movement of median price over time in txhousing. Color equal to city.
```{r}

```
## Exercise 4.6 Recreate the Mean v Median Scatterplot from the Abstract Question. Try to identify if any particular cities have an outlier problem. Visualize this in some way.
```{r}

```
## Exercise 4.7 Now do this for all years, not just 2015
```{r}

```




